---
title: "Bayesian Baseball"
output:
  pdf_document: default
  html_notebook: default
---

First, we load the necessary library packages.
```{r, echo=FALSE, include = FALSE}
library(tidyverse)
fill = "#4271AE"; line = "#1F3552"
```

Below is some exploratory data analysis.

I did some summary statistics on the data by position.
```{r}
ball_data = read.csv("BattingAverage.csv")

#mean statistics on hit rate of positions
ball_data %>%
  group_by(PriPos) %>%
  summarise(
    avg_atbats = mean(Hits),
    sd_atbats = sd(Hits),
    avg_hits = mean(AtBats),
    sd_hits = sd(AtBats),
    avg_hitrate = mean(Hits/AtBats),
    sd_hitrate = sd(Hits/AtBats)
)

#histograms show you the shape of distributions 
#boxplots help you compare between positions

#Standard deviations 

#count how many positions are in data
ball_data %>%
  count(PriPos)
```

```{r}
#count of the number of positions in the data
ggplot(data = ball_data) + geom_bar(mapping = aes(x = PriPos)) + ggtitle("Number of Positions in Data")
```
Compared to the other positions, there are much more pitcher data in the baseball dataset. 

```{r}
#distribution of average hit rate according to position 
ggplot(data = ball_data) + geom_boxplot(mapping = aes(x = PriPos, y = Hits/AtBats), fill= fill, color =  line) + ggtitle("Hit Rate of Each Position") + xlab("Position") +ylab("Hit Rate")
#one pitcher at 1 due to lack of atbats
```

```{r}
#distribution of at bats according to position
ggplot(data = ball_data) + geom_boxplot(mapping = aes(x = PriPos, y = AtBats)) + ggtitle("AtBats of each Position")
```

```{r}
#distribution of hits according to position
ggplot(data = ball_data) + geom_boxplot(mapping = aes(x = PriPos, y = Hits))
```
From the two graphs above, we can see that, compared with other positions, pitchers have a low hit rate due to their lack of chances atbats and low hits. 

```{r}
#comparison between at bats with hit rate
ggplot(data = ball_data) + geom_point(mapping = aes(x = AtBats, y = Hits/AtBats)) + facet_wrap(~PriPos)

```
Even when players have more chance at bats, their hit rate does not increase. 

```{r}
#correlation between data
dat_catcher = filter(ball_data, PriPos == "Catcher")
dat_leftfield = filter(ball_data, PriPos == "Left Field")
dat_shortstop = filter(ball_data, PriPos == "Shortstop")
rcorr(as.matrix(cbind(dat_catcher$Hits, dat_leftfield$Hits, dat_shortstop$Hits)))

#At the end make observations
```
Below is some code to help in analyzing patterns
```{r}
#find pitcher with zero hits
zero = filter(ball_data, Hits == 0 & PriPos == "Pitcher")
#find pitcher with 100% hit rate
perfect = filter(ball_data, Hits/AtBats == 1 & PriPos == "Pitcher")
```